Black-hole ringdown as a probe of higher-curvature gravity theories

### 2022-04-25 papers/paper_alt_theor_bounds/paper_alt_theor_bounds.tex
changes to qeft

Let us now consider the quartic EFT of GR, as our final example.
%
In Fig.~\ref{fig:qEFT_exec_sum} we show the posteriors on $\ell_{\rm qEFT}$ for GW150914 (top panel),
GW200129 (middle panel) for $N_{\rm max} = 0$, which are qualitatively similar to the cubic EFT of GR.
%
We find that both the EFT~\eqref{eq:eft_bound} and ParSpec~\eqref{eq:parspec_bound} are satisfied,
hence allowing us to infer bounds on $\ell_{\rm qEFT}$.
%
As Fig.~\ref{fig:qEFT_cdf} shows the CDFs for both events reach 90\% for $\gm = M_{\rm f}$ before $\varepsilon = 1/2$.
%
Consequently, from the joint posterior of $\ell_{\rm qEFT}$ we quote our final result,
%
\begin{equation}
    \ell_{\rm qEFT} \leqslant 51.3~\textrm{km} \quad \textrm{at 90\% credible level}\,.
\end{equation}
%
% \abcomm{Please update the paragraph below.}
In this theory, we have considered only $N_{\rm max} = 0$.
%
\ab{Indeed, we find that when adding the spin corrections and for $\ell_{\rm qEFT} \gtrsim 80$~km,
the waveforms can have a ringdown segment larger (sometimes seconds long) than the inspiral-plunge segment in the detectors frequency band, making the parameter estimation challenging.} \abcomm{The reference to 80 km comes out of the blue. Doesn't it? Why did you refer to it?}
% for which the ringdown portions of the larger than the inspiral-plunge part. \abcomm{The last sentence
% has some problems in English. I do not udnerstand it.}
%
To overcome this issue we have lowered the value of $\ell^{\rm max}_{\rm qEFT}$, but by doing so we have obtained posteriors
which were flat, just as our prior, and were thus uninformative. Hence, we do not quote any results for $N_{\rm max} = 1$.
%
Table~\ref{tab:summary_qeft} summarizes our findings for the quartic EFT of GR.
%
\begin{figure}[t]
\includegraphics[width=\columnwidth]{figs/qeft_posteriors_combined_corrected.pdf}
\caption{(Color online).~Posterior distribution function for the coupling constant $\ell_{\rm qEFT}$ for
GW150914 (top panel), GW200129 (middle panel) and combined events (bottom panel).
}
\label{fig:qEFT_exec_sum}
\end{figure}

\begin{figure}[t]
\includegraphics[width=\columnwidth]{figs/qeft_cdf_varying_threshold_corrected.pdf}
\caption{Similar to Fig.~\ref{fig:edgb_cdf}, but for the quartic EFT of GR.
We find that similarly to what happens in dCS gravity and the cubic EFT of GR, we
can place a constrain on $\ell_{\rm qEFT}$ when $\gm = M_{\rm f}$ for both events at 90\% credibility.
}
\label{fig:qEFT_cdf}
\end{figure}

@@

We presented an unified framework that combines the ParSpec framework
to model deviations to the GR QNMs~\cite{Maselli:2019mjd} with the \pSEOB~waveform model~\cite{Brito:2018rfr,Ghosh:2021mrv}.
%
We showed with concrete examples, how theory-specific QNM calculations of
slowly-rotating BHs in modified gravity theories can be mapped onto the
non-GR parameters of the ParSpec formalism.
%
The resulting \pSEOB~waveform model \ab{does} not bias (relative to GR) the inference of
the intrinsic binary parameters, as required by ParSpec (see, in particular, Fig.~\ref{fig:corner_plot} and
Fig.~\ref{fig:corner_plot_all} in Appendix~\ref{app:big_contour}),
%
Put together this allowed us to test four modified gravity theories (EdGB, dCS,
cubic, and quartic EFTs of GR) using observational data from the LVK events
GW150914 and GW200129. Our results are summarized in Table~\ref{tab:bound_summary}.

@@

For dCS gravity, we follow Ref.~\cite{Wagle:2021tam}, which numerically calculated the QNMs, $\sigma_{\rm dCS}$,
of slowly-rotating BHs, and found that for the axial gravitational modes the damping time increases, as we
increase the Chern-Simons coupling $\ell_{\rm dCS}$, \hs{at constant BH spin.}
%
Hence, according to the prescription of Sec.~\ref{sec:theory_specific_qnm},
this is the branch of QNMs we choose to work with.

@@

We proceed in the same away for the quartic EFT. Here we use Eq.~(68) (with
$\epsilon_{1} = 1$ and $\epsilon_{2} = 0$) and Eq.~(70) of
Ref.~\cite{Cano:2021myl},
%
In this case we find that both modes (``-'' and ``+'' in the notation of Ref.~\cite{Cano:2020cao}) make the
fundamental QNM damp fast, and we choose the ``-'' mode which decreases less the damping time. \abcomm{Not clear.
Decreases less with respect to what? Do you mean it has a smaller damping time?}
%
\ab{This time we} then find \sout{this time} that $p_{\rm qEFT} = 6$,
%
\begin{align}
    \delta \omega_{\rm qEFT}^{(0)} = -0.2114, \quad \delta \tau_{\rm qEFT}^{(0)} = -0.6070,
    \nonumber \\
    \delta \omega_{\rm qEFT}^{(1)} = -1.5263, \quad \delta \tau_{\rm qEFT}^{(1)} = 171.35.
\end{align}
%
As with the previous theories, the larger values of some of these parameters are compensated
by the assumptions of small coupling and small spin used to calculate the QNM frequencies.
\
\section{The estimation of  intrinsic binary parameters in General Relativity and modified theories of gravity}
\label{app:big_contour}

>>>>

Let us now consider the quartic EFT of GR, as our final example.
%
In Fig.~\ref{fig:qEFT_exec_sum} we show the posteriors on $\ell_{\rm qEFT}$ for GW150914 (top panel),
GW200129 (middle panel) for $N_{\rm max} = 0$, which are qualitatively similar to the cubic EFT of GR.
%
We find that while the Parspec bound is satisfied, the EFT bound is only marginally so,
%
As shown in Fig.~\ref{fig:qEFT_cdf}, the 90\% credible level is reached for $\varepsilon \approx 0.58$
(in the case of GW15094) and for $\varepsilon \approx 0.64$ (in the case of GW200129).
%
Having in mind that the hard cut-off $\varepsilon = 1/2$ is not fundamental, but to keep consistency
across our analysis, our final result
%
\begin{equation}
    \ell_{\rm qEFT} \leqslant 51.3~\textrm{km} \quad \textrm{at 90\% credible level},
\end{equation}
%
at 90\% credible level should be taken lightly.
%
However, we \emph{can} claim the validity of the bound above, but at a lower,
68\% credible level.

% We find that both the EFT~\eqref{eq:eft_bound} and ParSpec~\eqref{eq:parspec_bound} are satisfied,
% hence allowing us to infer bounds on $\ell_{\rm qEFT}$.
% %
% As Fig.~\ref{fig:qEFT_cdf} shows the CDFs for both events reach 90\% for $\gm = M_{\rm f}$ before $\varepsilon = 1/2$.
% %
% Consequently, from the joint posterior of $\ell_{\rm qEFT}$ we quote our final result,
% %
% \begin{equation}
%     \ell_{\rm qEFT} \leqslant 51.3~\textrm{km} \quad \textrm{at 90\% credible level}\,.
% \end{equation}
% %
% % \abcomm{Please update the paragraph below.}
In this theory, we have considered only $N_{\rm max} = 0$.
%
\ab{We find that the addition of spin corrections (while maintaining the same prior ranges on $\ell_{\rm qEFT}$ as
used in the $N_{\rm max} = 0$ study)
can result in waveforms that can have a ringdown segment larger (sometimes seconds long) than the inspiral-plunge segment in the detectors frequency band, making the parameter estimation challenging.} \abcomm{The reference to 80 km comes out of the blue. Doesn't it? Why did you refer to it?}~\hscomm{That was the upper prior limit for the GW150914 run. I modified the text to explain this.}
% for which the ringdown portions of the larger than the inspiral-plunge part. \abcomm{The last sentence
% has some problems in English. I do not udnerstand it.}
%
To overcome this issue we have lowered the value of $\ell^{\rm max}_{\rm qEFT}$, but by doing so we have obtained posteriors
which were flat, just as our prior, and were thus uninformative. Hence, we do not quote any results for $N_{\rm max} = 1$.
%
Table~\ref{tab:summary_qeft} summarizes our findings for the quartic EFT of GR.
%
\begin{figure}[t]
\includegraphics[width=\columnwidth]{figs/qeft_posteriors_combined_corrected.pdf}
\caption{(Color online).~Posterior distribution function for the coupling constant $\ell_{\rm qEFT}$ for
GW150914 (top panel), GW200129 (middle panel) and combined events (bottom panel).
}
\label{fig:qEFT_exec_sum}
\end{figure}

\begin{figure}[t]
\includegraphics[width=\columnwidth]{figs/qeft_cdf_varying_threshold_corrected.pdf}
\caption{Similar to Fig.~\ref{fig:edgb_cdf}, but for the quartic EFT of GR.
We find that similarly to what happens in dCS gravity and the cubic EFT of GR, we
can place a constrain on $\ell_{\rm qEFT}$ when $\gm = M_{\rm f}$ for both events at \hs{68\% credibility}.
}
\label{fig:qEFT_cdf}
\end{figure}

@@

We presented an unified framework that combines the ParSpec framework
to model deviations to the GR QNMs~\cite{Maselli:2019mjd} with the \pSEOB~waveform model~\cite{Brito:2018rfr,Ghosh:2021mrv}.
%
We showed with concrete examples, how theory-specific QNM calculations of
slowly-rotating BHs in modified gravity theories can be mapped onto the
non-GR parameters of the ParSpec formalism.
%
The resulting \pSEOB~waveform model does not bias (relative to GR) the inference of
the intrinsic binary parameters, as required by ParSpec (see, in particular, Fig.~\ref{fig:corner_plot} and
Fig.~\ref{fig:corner_plot_all} in Appendix~\ref{app:big_contour}),
%
Put together this allowed us to test four modified gravity theories (EdGB, dCS,
cubic, and quartic EFTs of GR) using observational data from the LVK events
GW150914 and GW200129. Our results are summarized in Table~\ref{tab:bound_summary}.

@@

For dCS gravity, we follow Ref.~\cite{Wagle:2021tam}, which numerically calculated the QNMs, $\sigma_{\rm dCS}$,
of slowly-rotating BHs, and found that for the axial gravitational modes the damping time increases, as we
increase $\ell_{\rm dCS}$, \hs{at constant BH spin.}
%
Hence, according to the prescription of Sec.~\ref{sec:theory_specific_qnm},
this is the branch of QNMs we choose to work with.

@@

We proceed in the same away for the quartic EFT. Here we use Eq.~(68) (with
$\epsilon_{1} = 1$ and $\epsilon_{2} = 0$) and Eq.~(70) of
Ref.~\cite{Cano:2021myl},
%
In this case we find that both modes (``-'' and ``+'' in the notation of Ref.~\cite{Cano:2020cao}) make the
fundamental QNM damp fast, and we choose the ``-'' mode which decreases less the damping time. \abcomm{Not clear.
Decreases less with respect to what? Do you mean it has a smaller damping time?}
%
\ab{This time we} then find \sout{this time} that $p_{\rm qEFT} = 6$,
%
\begin{align}
    \delta \omega_{\rm qEFT}^{(0)} = -0.2114, \quad \delta \tau_{\rm qEFT}^{(0)} = -0.6070,
    \nonumber \\
    \delta \omega_{\rm qEFT}^{(1)} = -1.5263, \quad \delta \tau_{\rm qEFT}^{(1)} = 171.35.
\end{align}
%
As with the previous theories, the larger values of some of these parameters are compensated
by the assumptions of weak coupling and small spin used to calculate the QNM frequencies.
\
\section{The estimation of  intrinsic binary parameters in General Relativity and modified theories of gravity}
\label{app:big_contour}
